Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2023-04-03 arxiv/main.tex
respond to comments

Whatever the cause, global biosphere disturbance theory represents an alternative paradigm in disease spreading dynamics. Our goal with this study is to present a novel mathematical bridge between these two paradigms which at its core is a connection between the logistic growth model and the Gompertzian growth model. This connection represents an alternative to the well-known parametrization given by Richard (1959) \cite{richards1959flexible}, where he augmented standard logistical growth by one parameter. This is why it is also sometimes called $\theta$-logistic growth \cite{petroni2020logistic}.
% \begin{equation}
% \frac{d}{dt}\ln{Y(t)} = -\beta\left[1 - \left(\frac{Y(t)}{{Y_{\infty}}}\right)^\theta\right],
% \end{equation}
% where traditional logistic growth is augmented by a parameter $\theta \in \mathbb R^+$.

 One issue with Richard's parametrization is that the traditional communicable disease models cannot be realistically extrapolated to the Gompertz model, as shown by Wang et \textit{al.} (2012) \cite{wang2012richards}. The reason for why Richard's model cannot realistically represent a communicable disease with a Gompertz model is because it would imply that there is ``long-range interaction'' between the susceptible and infected (see \cite{petroni2020logistic} for details), which is counter to the basic assumption that communicable diseases spread through proximity between the susceptible and infected. Based on these findings, one would then be forced to conclude that Gompertz growth cannot arise in the context of a communicable disease.

>>>>

Whatever the cause, global biosphere disturbance theory represents an alternative paradigm compared to the traditional contact-transmissible contagion theory in the context of disease spreading dynamics. Our goal with this study is to present a novel mathematical bridge between these two paradigms which at its core is a connection between the Gompertzian growth model and the logistic growth model. Our proposed bridge represents an alternative to the well-known parametrization given by Richard (1959) \cite{richards1959flexible}, where he augmented standard logistical growth by one parameter. This is why it is also sometimes called $\theta$-logistic growth \cite{petroni2020logistic}.
% \begin{equation}
% \frac{d}{dt}\ln{Y(t)} = -\beta\left[1 - \left(\frac{Y(t)}{{Y_{\infty}}}\right)^\theta\right],
% \end{equation}
% where traditional logistic growth is augmented by a parameter $\theta \in \mathbb R^+$.

 One issue with Richard's parametrization is that the traditional communicable disease models cannot be realistically extrapolated to the Gompertz model, as shown by Wang et \textit{al.} (2012) \cite{wang2012richards}. The reason why Richard's model cannot realistically represent a communicable disease with a Gompertz model is because it would imply that there is ``long-range interaction'' between the susceptible and infected (see \cite{petroni2020logistic} for details), which is counter to the basic assumption that communicable diseases spread through proximity between the susceptible and infected. Based on these findings, one would then be forced to conclude that Gompertz growth cannot arise in the context of spread through proximity between the susceptible and infected.
